Gsmodutils: A python based framework for test driven genome scale metabolic model development

### 2017-11-29 article/article.tex
Expanded article to cover c aceto examples

\section{Example usage: Clostridium acetobutylicum models}

% Scientific background

% Prodblems models solve

% Created designs in gsmodutils - start with original model, make changes

% Diffs between existing models

% Conclusions

@@

>>>>

\section{Example usage Clostridium acetobutylicum models}
\subsection{Scientific background}
\textit{C. acetobutylicum} ATCC 824 is a widely researched organism that has found use in the production of solvents for decades.
More recently, \textit{C. acetobutylicum} has been applied for the production of alcohols such as ethanol and n-butanols.
The industrial relevance and strong background of experimental data has led to the application of many genome scale models \cite{}.
Whilst increasingly informative when combined with experimental data, the history of the management of models for \textit{C. acetobutylicum} tells a different story.

Initially, two rival models were published containing a similar number of genes \cite{lee2008genome, senger2008genome}.
% \textbf{TODO: describe in more detail}.
%Discuss limitations of rival models
This led to the further development of subsequent models, iCaC 490 \cite{mcanulty2012genome} and an automatically generated model from  the Model SEED  pipeline \cite{}.

In this example we focus on the more recent contributions of iCac802 \cite{dash2014capturing} and iCac967 \cite{yoo2015quantitative}.
iCac802 combines a significantly updated set of annotations from previously curated model versions, an improved TCA cycle, full proton balancing in reactions and reactions for more complete coverage of hexadecanoyl-acp, hexadecanoyl-CoA, crotonyl-CoA pathways as well as pathways for the uptake and biosynthesis of more compounds.
Crucially, the model itself received significant validation against fluxonomic data from 13C-metabolic flux analysis ($13$C-MFA).
This additional validation set considerable flux bounds upon pathways, leading to improved curation.

Whilst iCac802 was informed by previous models, critically, it was started as a new curation project, a significant undertaking of work.
Such an approach may be necessary to uncover errors or oversights in construction processes.
However, this naturally adds a significant time burden for model curators and may be due to a lack of adaptability from previous models.

The iCac 967 model presents a smaller set of changes to the annotation of the \textit{C. acetobutylicum} model.
While the number of genes increased from 802 to 967 manual annotation actually reduced the number of reactions and metabolites from 1462 to 1137 and 1231 to 1058, respectively.
The annotation ane experimental evidence from iCac802 \cite{dash2014capturing} to iCac967 \cite{yoo2015quantitative} is greatly expanded.
However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC907 model preserves crucial experimental validation.
The following sections aim to check the later model against previous validation criteria.

\subsection{Model integration}
\textbf{Project construction:}
An initial project of the \textit{C. acetobutylicum} model was created in gsmodutils.
This is based around the iCac802 base model.
Included are also knockout strains $\Delta adc, \Delta ack, \Delta ptb$ and $\Delta hbd$ (TODO: describe knockouts).

In addition, a design listing the constraints changed in model version iCaC907 
\footnote{We would like to note that such a significant change to a model, in many practical cases, would not be considered a design but, rather, a separate model.
Separate models within the same project is a feature supported by gsmodutils.}.
\\
\textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.

\subsection{Evaluation of model validation criteria}
Several crucial validation criteria were identified from the experimental and theoretical analysis of the models found in the original publications.
We note that this is not an exhaustive set of criteria.
This section is designed to serve as a use case for the gsmodutils framework, not to comprehensively review the initial publications.
\\
\textbf{TCA cycle}\\
\textbf{Proton balanced results}\\
\textbf{Predictions of knockouts}\\
\textbf{Flux validation against 13C data}\\
\textbf{Under acidogenic stress (H inhibited conditions) } \\
\textbf{Solvent formation Cell recycling in solventogenic phase}\\

% Conclusions
\subsection{Remarks} 
\textbf{To fill in when validation is complete.}

@@

% Furthermore, several methods for the automated construction of gsms exist.
% However, to our knowledge no methods include a fully automated pipeline for updating existing manually created models once constructed.
% We would like to note that the collection of data for the validation of models and the construction of tests could be used to allow automated updates to take place.
% For example, as GPR are curated models could automatically be updated with the inclusion of new information whilst being validated against existing criteria.
